---
title: "Random Forest Final Model"
author: ""
date: ""
output: html_document
---

```{r setup1, include = FALSE}
def.chunk.hook <- knitr::knit_hooks$get("chunk")
knitr::opts_chunk$set(cache = FALSE)
knitr::knit_hooks$set(
  chunk = function(x, options) {
    x <- def.chunk.hook(x, options)
    ifelse(options$size != "normalsize", paste0("\n \\", options$size, "\n\n", x, "\n\n \\normalsize"), x)
  }
)
# knitr::knit_hooks$set(inline = function(x) {
#   prettyNum(round(x, 2), big.mark = ",")
# })
options(scipen=999)
```

\begin{center}
\Huge{PPOL 670 | Final Project}

\Huge{Random Forest}
\end{center}

\vspace{0.1in}

[GitHub](https://github.com/elenabagnera/CapitalBikesProject)

```{r setup, include=FALSE}
library(tidyverse)
library(tidymodels)
library(recipes)
library(vip)
library(lubridate)
library(workflows)
library(ranger)
source("times.R")
source("io.R")
source("manipulate_data.R")
doParallel::registerDoParallel()
```

## Preparing the Data

```{r get_data, cache=TRUE, warning=FALSE, message=FALSE}

lagged_data <- read_csv('lagged_data_3200.csv')
hour_data <- read_csv('hour_data_3200.csv')

model_data <- lagged_data %>%
  add_predictor_times() %>%
  filter(!is.na(departures) & !is.na(arrivals)) %>%
  # Having real time arrival data would be cheating
  select(-arrivals) %>% 
  format_weather()
```


```{r setup_data}

set.seed(28021995)

# create a split object
data_split <- initial_split(model_data, prop = 0.8)

# create the training and testing data
train <- training(x = data_split)
test  <- testing(x = data_split)
```

## Our preprocessing for Random Forests is the same as decision trees because Random Forests is just a "more complex" decision tree.

```{r recipe, cache=TRUE}
recipe <-
  recipe(formula = departures ~ ., data = train) %>%
  step_holiday(date, holidays = timeDate::listHolidays("US"), keep_original_cols = FALSE) %>% 
  step_nzv(all_predictors()) 
  # step_BoxCox(all_outcomes()) # Does not run with zero in the values
```

## Model Implementation

```{r tuning, cache=TRUE}
# Tuning min_n doesn't seem to help
mod <- rand_forest(mtry = 250, trees = 1000) %>%
  set_mode("regression") %>%
  set_engine("ranger")

wf <- workflow() %>%
  add_recipe(recipe) %>%
  add_model(mod)
```


```{r chosen_one, cache=TRUE}
final_wf <- workflow() %>%
  add_recipe(recipe) %>%
  add_model(mod) %>%
  fit(train)

p_test <- bind_cols(
  test,
  predict(object = final_wf, new_data = test))
```

## Results

Out of sample RMSE

```{r OOS_RMSE_lincoln}
sqrt(mean((p_test$departures - p_test$.pred)^2))
write_csv(p_test, "final-p-test.csv")
```

```{r lincoln_final_vip, warning=FALSE, eval=FALSE}
mod %>%
  set_engine("ranger", importance = "permutation") %>%
  fit(departures ~ .,
    data = train
  ) %>%
  vip(geom = "point")
```


## Discussion of Results

### Average departure per hour

```{r mean_of_departures}
p_test %>% summarize(mean(departures))
```

### Top 20 hours of 12431 hours.

```{r top_20_departures}
p_test %>% arrange(-departures) %>% select(departures) %>% head(20)
```

### Top 20 Error - very high departure days

```{r top_20_error}
p_test %>% 
  mutate(error = abs(.pred - departures)) %>% 
  select(year, month, day, hour, weekday, departures, .pred, error, conditions,temperature, cloud_cover, snow, precipitation, lincoln_memorial_departures) %>% 
  arrange(-error) %>% 
  head(20)
```

### Best guesses

```{r}
p_test %>% 
  mutate(error = abs(.pred - departures)) %>% 
  select(year, month, day, hour, weekday, departures, .pred, error, conditions,temperature, cloud_cover, snow, precipitation, lincoln_memorial_departures)  %>% 
  arrange(error) %>% 
  head(15)
```

### Best guesses with more 5 departures

```{r}
p_test %>% 
  mutate(error = abs(.pred - departures)) %>% 
  filter(departures > 5) %>% 
  select(year, month, day, hour, weekday, departures, .pred, error, conditions,temperature, cloud_cover, snow, precipitation, lincoln_memorial_departures)  %>% 
  arrange(error) %>% 
  head(15)
```

### Worst guessing with low departures - why?

```{r}
p_test %>% 
  mutate(error = abs(.pred - departures)) %>% 
  filter(departures < 5) %>% 
  select(year, month, day, hour, weekday, departures, .pred, error, conditions,temperature, cloud_cover, snow, precipitation, lincoln_memorial_departures)  %>% 
  arrange(-error) %>% 
  head(15)
```

### Worst guesses with low departures and actual weather

```{r}
p_test_no_wx <- p_test %>% 
    mutate(error = abs(.pred - departures)) %>% 
  select(year, month, day, hour, weekday, departures, .pred, error, lincoln_memorial_departures)
  
hour_data_wx <- hour_data %>% 
  select(year, month, day, hour, conditions,temperature, cloud_cover, precipitation, wind_speed, heat_index, wind_chill)

left_join(p_test_no_wx, hour_data_wx, on=c(year, month, day, hour)) %>% 
  filter(departures < 5) %>% 
  arrange(-error) %>% 
  head(15)

p_test %>% summarize(mean(wind_speed))
```

### Can we predict Lincoln Memorial will be empty? Every prediction for Sunday.

```{r}
p_test %>% 
  filter(weekday == "Sun") %>% 
  ggplot(aes(x = hour, y = .pred))+
  geom_point(color = "orange")
  # geom_smooth()
```

### After testing multiple variables, two week lagged data from two stations heavily weighted departures

```{r}
p_test %>% 
  filter(weekday == "Sun", hour == 14) %>% 
  select(`4th_st_&_madison_dr_nw_arrivals`, `4th_st_&_madison_dr_nw_departures`, `lincoln_memorial_departures`, lincoln_memorial_arrivals, .pred) %>% 
  pivot_longer(1:4, names_to = "station", values_to="value") %>% 
  ggplot(aes(x = value, y = .pred, color=station))+
  geom_point()
```

### Filtering out the high data point (Daylight Savings Time)

```{r}
# Only having one data point for certain times of the year hurts
# That one high datapoint is daylight savings time....
# only 23 hours in that day
p_test %>% 
  filter(weekday == "Sun", hour == 14) %>% 
  select(`4th_st_&_madison_dr_nw_arrivals`, `4th_st_&_madison_dr_nw_departures`, `lincoln_memorial_departures`, lincoln_memorial_arrivals, .pred) %>% 
  pivot_longer(1:4, names_to = "station", values_to="value") %>% 
  filter(value >= 1) %>% 
  ggplot(aes(x = value, y = .pred, color=station))+
  geom_smooth(aes(x = value, y = .pred, color=station), alpha=0.1)+
  geom_point()

# Look at Georgetown Harbor / 30th St NW
# 39th & Calvert St NW / Stoddert
```

Having two or more years would likely have helped that did not have a large change due to COVID.

## High Supply

```{r high_supply, cache=TRUE}
cbs_data <- read_files('data/')
data <- sep_departures_from_arrivals(cbs_data) %>%
  filter(!is.na(station)) # Remove electric bikes not going to / from a station

filtered_data <- filter_by_distance(data, from_station = 'georgetown_harbor_/_30th_st_nw', distance_m = 1600)

hour_data <- get_station_hourly(filtered_data) %>%
  get_historic_weather() %>%
  add_sun_is_out() %>%

lagged_data <- setup_for_time_prediction(hour_data, station_name = "georgetown_harbor_/_30th_st_nw")

model_data <- lagged_data %>%
  add_predictor_times() %>%
  filter(!is.na(departures) & !is.na(arrivals)) %>%
  select(-arrivals) %>% 
  format_weather()
```

```{r chosen_one_supply, cache=TRUE}
final_wf <- workflow() %>%
  add_recipe(recipe) %>%
  add_model(mod) %>%
  fit(train)

p_test <- bind_cols(
  test,
  predict(object = final_wf, new_data = test))
```

## High Supply Results

Out of sample RMSE

```{r OOS_RMSE_Supply}
sqrt(mean((p_test$departures - p_test$.pred)^2))
```

```{r supply_final_vip, warning=FALSE, eval=FALSE}
mod %>%
  set_engine("ranger", importance = "permutation") %>%
  fit(departures ~ .,
    data = train
  ) %>%
  vip(geom = "point")
```

```{r high_supply_day}
p_test %>% 
  filter(weekday == "Sun") %>% 
  ggplot(aes(x = hour, y = .pred))+
  geom_point(color = "orange")
  # geom_smooth()
```

## High Demand

```{r high_demand, cache=TRUE}
filtered_data <- filter_by_distance(data, from_station = '39th_&_calvert_st_nw_/_stoddert', distance_m = 1600)

hour_data <- get_station_hourly(filtered_data) %>%
  get_historic_weather() %>%
  add_sun_is_out() %>%

lagged_data <- setup_for_time_prediction(hour_data, station_name = "39th_&_calvert_st_nw_/_stoddert")

model_data <- lagged_data %>%
  add_predictor_times() %>%
  filter(!is.na(departures) & !is.na(arrivals)) %>%
  select(-arrivals) %>% 
  format_weather()
```

```{r chosen_one_demand, cache=TRUE}
final_wf <- workflow() %>%
  add_recipe(recipe) %>%
  add_model(mod) %>%
  fit(train)

p_test <- bind_cols(
  test,
  predict(object = final_wf, new_data = test))
```

## Results

Out of sample RMSE

```{r OOS_RMSE_Demand}
sqrt(mean((p_test$departures - p_test$.pred)^2))
```

```{r demand_final_vip, warning=FALSE, eval=FALSE}
mod %>%
  set_engine("ranger", importance = "permutation") %>%
  fit(departures ~ .,
    data = train
  ) %>%
  vip(geom = "point")
```

```{r high_demand_day}
p_test %>% 
  filter(weekday == "Sun") %>% 
  ggplot(aes(x = hour, y = .pred))+
  geom_point(color = "orange")
  # geom_smooth()
```

